Redundant calibration as a complex optimization problem

### 2016-12-06 paper.tex
added some text to sparsity section

[Redundant calibration has been in use for a long time. It was already being used in the 1980's. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
LOGCAL is reviewed in Appendix ??. The next improvement was to perform a firt-order taylor expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
perturbations of the model-parameters. This approach is commonly referred to as LINCAL. LINCAL is derived from first principals in Appendix . Two ways of speeding up redundant calibration has also been 
proposed. The first technique makes use of the alternating direction implicit method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
and in doing so speedup the algorithm itself was proposed in . A proper comparison between these two approaches has not been done before.

@@

\subsubsection{Spectral Condition Number}
kappa + itr plots
\subsubsection{Sparsity}
sparsity derivation
sparsity + computational complexity plots

>>>>

[Redundant calibration has been in use for a long time. Noordam used redundant calibration to produce high-dynamic range images of 3C84 with the 
WSRT array.] RW The first implementations of redundant calibration made use of the logarithm function, because when the logarithm is applied to the scalar measurement equation redundant 
calibration is transformed into a linear problem cite Wieringa,camps and Lui. This approach is also generally referred to as LOGCAL.
LOGCAL is reviewed in Appendix ??. The next improvement was to perform a firt-order taylor expansion (which is equivalent to the GN algorithm Kurien 2016) of the measurement equation an then to solve for the
perturbations of the model-parameters. This approach is commonly referred to as LINCAL. LINCAL is derived from first principals in Appendix . Two ways of speeding up redundant calibration has also been 
proposed. The first technique makes use of the alternating direction implicit method, in which we alternate between solving the gains, their conjugates and 
the true observed visibilities (wijnholds and marthi). Using the conjugate gradient method to exploit the sparsity inherent in the redundant calibration problem, 
and in doing so speedup the algorithm itself was proposed in . A proper comparison between these two approaches has not been done before.

@@

\subsubsection{Spectral Condition Number}
kappa + itr plots
\subsubsection{Sparsity}

% \begin{enumerate}
% \item $\boldsymbol{H}$: is a $(4N-2)\times(4N-2)$ or a $P \times P$ matrix. $N$ denotes the number of antennas and $P$ denotes the number of parameters. This matrix
% contains $6N^2 - 2N - 2$ non-zero entries. It contains $16N^2 - 16N + 4$ entries. We can construct this matrix with $4N^2 -  4N$ or $\frac{1}{4} P^2 -1$ elementary operations.
% \item $\boldsymbol{A}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $N^2 + N - 1$ non-zero entries. We can construct this matrix with $2\frac{1}{2} (N^2 -  N)$ elementary operations.  
% \item $\boldsymbol{B}$: is a $(2N-1)\times(2N-1)$ matrix. This matrix has $2N^2 - 2N$ non-zero entries. We can construct this matrix with $1\frac{1}{2} (N^2 -  N)$ elementary operations.
% \item $\boldsymbol{C}$: is a $N\times N$ matrix. This matrix has $N$ non-zero entries. We can construct this matrix with $N^2-N$ elementary operations.
% \item $\boldsymbol{D}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations. 
% \item $\boldsymbol{E}$: is a $(N-1) \times (N-1)$ matrix. This matrix has $(N-1)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.  
% \item $\boldsymbol{F}$: is a $N \times N$ matrix. This matrix has $N^2 - N$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
% \item $\boldsymbol{G}$: is a $N \times (N-1)$ matrix. This matrix has $\frac{1}{2} (N^2 -  N)$ non-zero entries. We can construct this matrix with $\frac{1}{2} (N^2 -  N)$ elementary operations.
% \item $\boldsymbol{0}$: is a $(N-1) \times (N-1)$ all zero matrix. 
% \end{enumerate}

%We can construct $\boldsymbol{C}-\boldsymbol{G}$ in $O(N^2)$

As alluded to in section ..., the second factor that determines the computational complexity of the conjugate gradient method when used to perform redundant calibration is the amount of non-zero
entries in the Hessian matrix $\bH$, which is the matrix we are trying to invert. We have also shown in section that the first factor, the spectral condition number of $\bH$, can be reduced 
to a constant by employing preconditioning and can therefore be completely eliminated from the computational complexity equation. A useful matrix quantity
which we will use in this section is the sparsity factor $\gamma$ of a matrix, which is the ratio between the amount of zero entries and the total amount of entries in a matrix. We 
can therefore define $\gamma$ as follows:
\begin{equation}
 \gamma = \left (1 - \frac{m}{P^2} \right ) 
\end{equation}
It is trivial to determine an analytic expression for the sparsity ratio of $\bH$ when our redundant array is in a regular east-west geometric configuration
The sparsity and the asymptotic sparsity ratio of $\boldsymbol{H}$ are equal to
\begin{align}
\gamma &= \frac{5N^2-7N+3}{8N^2-8N+2} & \gamma_{\infty} &= \lim_{N\rightarrow \infty}\gamma = \frac{5}{8}. 
\end{align}
However when our redundant array makes use of a more complicated geometric layout it becomes much more difficult to construct such expressions as $\bH$ itself 
becomes much harder to express analytically.

In Fig... we plot the sparsity ratio of three different geometric layouts as a function of the number antennas in the array. Discuss the solid lines 
and how they relate to the two equations from above.

We can now use the sparsity factor to compute the order of the computational cost of inverting $\bH$ as follows:
\begin{align}
P^{c} &= (1 - \gamma)P^2\\
c &= \log_{P}(1 - \gamma) + 2 & c_{\infty} &= \lim_{N\rightarrow \infty} c = 2
\end{align}

Asymptotically the computational cost of inverting $\bH$ is $O(P^2)$. However, the computational cost converges 
very slowly to its asymptotic value, and in general is equal to $O(P^{c})$, where $c \sim 1.7$ if a Hexagonal geometric layout is employed.

sparsity derivation
sparsity + computational complexity plots
